Heuristics for the Score-Constrained Strip-Packing Problem

### 2018-07-13 PAPER/main.tex
made changes to section 2, table

\begin{definition}
	\label{defn:althamcycle}
	Let $G(V, B\cup R)$ be an undirected graph, where each edge is a member of one of two sets, $B$ or $R$. $G$ contains an alternating Hamiltonian cycle if there exists a Hamiltonian cycle such that successive edges alternate between sets $B$ and $R$.
\end{definition}

\noindent It is clear than an alternating Hamiltonian cycle on $G$, if one exists, correpsonds to a feasible solution, as each ``blue'' edge from $B$ represents a pair of values in $\mathcal{M}$, and each ``red'' edge from $R$ indicates the values that can be placed alongside one another feasibly. Also, note that every edge in $B$ must be present in the alternating Hamiltonian cycle. Consequently, the task can also be seen as finding a matching $R' \subseteq R$ of cardinality $n+1$ such that the edge sets $B$ and $R'$ form an alternating Hamiltonian cycle as described in Definition \ref{defn:althamcycle}. The problem of finding an alternating Hamiltonian cycle on general graphs is NP-hard as it generalises the classical Hamiltonian cycle problem \cite{haggkvist1977}. However, the special structure of graphs derived from instances of the COP allows them to be solved in polynomial-time using the following method.

To find a matching $R' \subseteq R$, the Maximum Cardinality Matching (MCM) algorithm is executed, which is based on \cite{becker2010} and also the earlier work of Mahadev and Peled \cite{mahadev1994}. First, take each vertex $v_1, v_2,..., v_{2n+2}$ in turn, and select the edge in $R$ connecting $v_i$ to the highest-indexed vertex $v_j$ that is not already incident to an edge in $R'$. Now add this edge $(v_i, v_j)$ to $R'$, and proceed to the next vertex until all the vertices have been assessed. The two vertices incident to each matching edge in $R'$ are now referred to as being ``matched''. Similarly to partners, let $m : V \to V$ be a bijective function that assigns each vertex with its match, $m(v_i) = v_j$. Then, we can denote this matching set as $R' =\{(v_i, m(v_i)) : v_i \in V\}$.

During MCM, if a vertex $v_i$ is not adjacent to any other unmatched vertex except its partner $p(v_i)$ via a blue edge, the preceeding vertex $v_{i-1}$ can be ``rematched'', provided that 
\begin{enumerate*}[label={(\alph*)}]
	\item $v_i$ is not the first vertex; 
	\item the previous vertex $v_{i-1}$ has been matched; and
	\item $v_{i-1}$ and $p(v_i)$ are adjacent via a red edge in $R$.
\end{enumerate*}
Then, $v_i$ is matched with the vertex that is currently matched with $v_{i-1}$, and $v_{i-1}$ is rematched with $p(v_i)$. The complexity of MCM is $O(n \lg n)$ time \cite{mahadev1994}.

Clearly, if $|R'| < n+1$ after MCM has completed, there are insufficient red edges to form an alternating Hamiltonian cycle, and so no feasible solution exists for the given instance $\mathcal{M}$. Otherwise, $R'$ is a perfect matching of cardinality $n+1$ and the spanning subgraph $G'=(V, B \cup R')$ is a 2-regular graph, where each vertex $v_i \in V$ is incident to one blue edge and one red edge, as can be seen in Figure \ref{fig:mps}. $G'$ therefore consists of one or more cyclic components $C_1, C_2, ..., C_l$. In the case where $G'$ comprises one component, i.e. $l = 1$, then this alternating cycle is in fact Hamiltonian, and therefore specifies a solution to the COP. However, if $l > 1$, the components of $G'$ must be combined to form a single alternating Hamiltonian cycle. To do this, an additional procedure called Bridge Recognition (BR) is executed that selects suitable edges in $R\backslash R'$ to replace edges in $R'$ to connect the components of $G'$.

@@

\noindent BR operates by first ordering the edges in $R'$ such that the lower-indexed vertices of each edge are in increasing order and the higher-indexed vertices are in decreasing order. Any edges that cannot be placed in such an order are then removed from this list. For instance, in the example in Figure \ref{fig:mps}, the edges would be sorted as follows: $(v_1, v_{14}), (v_2, v_{13}),(v_3, v_{12}),(v_4, v_{11}),(v_6, v_{10}),(v_7, v_8)$. Note that, since $v_5$ was not matched with the highest-indexed vertex possible during MCM, the edge $(v_5, v_9)$ does not adhere to the required structure of the list and is therefore omitted.

@@

To find these overlapping sets, a Modified Bridge Recognition (MBR) algorithm is used, which is based on BR. Firstly, a copy of the set $R_i$ with the highest cardinality\footnote{In the event of a tie, MBR selects the set with the lowest index.} generated by BR is created, called $R^{*}_1$, and is added to $\mathcal{R}^{*}$. Then, MBR takes the original sorted list used in BR and removes edges from the list that are already in $\mathcal{R}^{*}$. Similarly to BR, MBR proceeds through the list to find an edge that meets the following conditions:
\begin{enumerate*}[label={(\alph*)}]
	\item The lower-indexed vertex of the current edge and the higher-indexed vertex of the next edge in the list are adjacent via an edge in $R$; and
	\item the current edge and an edge in $\mathcal{R}^{*}$ are members of the same component of $G'$, and the next edge is a member of a component that does not have an edge in $\mathcal{R}^{*}$, or vice versa.
\end{enumerate*}
If both conditions hold, the current edge is added to a new set $R^{*}_2$ which is then added to $\mathcal{R}^{*}$. MBR continues to add suceeding edges to $R^{*}_2$, provided condition (a) holds and the suceeding edge is a member of a component that does not have an edge in $\mathcal{R}^{*}$, stopping when an edge that does not meet these conditions is reached. Then, if every component of $G'$ has an edge in one of the sets in $\mathcal{R}^{*}$, these sets are able to connect all the components of $G'$ together, and so MBR terminates. Otherwise, the edges in $R^{*}_2$ are removed from the sorted list, and MBR repeats the search for suitable edges to start a new set $R^{*}_3$. This procedure continues until either $\mathcal{R}^{*}$ contains overlapping sets that cover all components of $G'$, or until there are no more suitable edges in the list to start a new set. If MBR has produced a feasible collection of sets then the components of $G'$ can be merged to create an alternating Hamiltonian cycle by applying the connecting procedure above to every set in $\mathcal{R}^{*}$.

Using the instance illustrated in Figure \ref{fig:connect/cycle} as an example, the edges $(v_3, v_{12})$ and $(v_4, v_{11})$ are in the set $R_1$ formed by BR, as \begin{enumerate*}[label={(\alph*)}]
	\item $(v_3, v_{11}) \in R$, i.e. the lower-indexed vertex of the first edge is adjacent to the higher-indexed vertex of the next edge; and
	\item the edges are members of different components ($(v_3, v_{12}) \in C_1$ and $(v_4, v_{11}) \in C_2$ as shown previously in Figure \ref{fig:mps}).
\end{enumerate*}
Note that since $v_4$ and $v_{11}$ are adjacent, $v_4$ must also be adjacent to $v_{12}$, i.e. $(v_4, v_{12}) \in R$, since the value associated with $v_{12}$ is greater than or equal to the value associated with $v_{11}$. Then, because $|R_1| = l$, the edges $(v_3, v_{12})$ and $(v_4, v_{11})$ are removed from $R'$ and replaced by the edges $(v_3, v_{11})$ and $(v_4, v_{12})$ from $R_1$, as illustrated in Figure \ref{fig:connect/cycle}. Removing the dominating vertices and any incident edges results in an alternating Hamiltonian path, which corresponds to a feasible solution $\mathcal{T}$.

@@

\section{Experimental Results}
For our experiments, we generated 1000 problem instances for $|\mathcal{I}| \in \{500, 100\}$. For each problem instance the items have varying widths $w_i$ between 150 and 1000 and score widths $a_i, b_i$ between 1 and 70 to ensure that each item has exactly two score lines, all selected randomly from a uniform distribution. Strips of widths $W = 5000$, 2500 and 1250 were used to influence the number of items per strip. As the width of the strips decrease, the average number of items per strip also decreases, making the problem more constrained. Both the items and the strips have equal height of $H = 1$. Similarly to experiments performed in \cite{lewis2011}, we also introduced a parameter $\delta$ to denote the proportion of pairs of score widths from different items that meet the vicinal sum constraint, i.e. whose sum is greater than or equal to $\tau$. Values of $\delta$ from 0.0 to 1.0 were created by changing the value of $\tau$. Clearly, when $\delta = 0.0$, there are no items that can be packed together feasibly, and so $n$ strips will be required (one for each item), whereas if $\delta = 1.0$ all pairs of score widths meet the constraint, and the problem is equivalent to the BPP.

The heuristics were implemented in C++ and executed on a computer with an Intel Core i3-2120 3.30GHz processor. Our source code and all data is provided at \cite{hawa2018}. Since optimal solutions are not available, in our case solution quality $q$ is estimated by comparing to the theoretical minimum, $q = |\mathcal{S}| / t$. For each heuristics, we calculated the average solution quality for every combination of $n$, $\delta$, and $W$ from 1000 instances. All individual trials were seen to complete in under 160ms.

\begin{table}[h!]
	\centering
	\caption{Average solution quality $q$ for $n = 500$.}
	\begin{tabular}{@{}c@{\hspace{15pt}}c@{\hspace{7pt}}c@{\hspace{5pt}}ccc@{\hspace{7pt}}c@{\hspace{5pt}}ccc@{\hspace{7pt}}c@{\hspace{5pt}}c@{}}\toprule
		& \multicolumn{3}{c}{$W = 5000$, $t = 58.039$} && \multicolumn{3}{c}{$W = 2500$, $t = 115.571$} && \multicolumn{3}{c}{$W = 1250$, $t = 230.648$} \\
		\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12}
		$\delta$ & MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$\\ \midrule
		0.0 & 8.618 & 8.618 & 8.618 && 4.328 & 4.328 & 4.327 && 2.169 & 2.169 & 2.169 \\
		0.1 & 5.214 & \textbf{4.842} & 5.161 && 2.659 & \textbf{2.477} & 2.657 && 1.515 & \textbf{1.479} & 1.515 \\
		0.2 & 4.031 & \textbf{3.459} & 3.976 && 2.121 & \textbf{1.847} & 2.118 && 1.326 & \textbf{1.318} & 1.326 \\
		0.3 & 3.111 & \textbf{2.348} & 3.038 && 1.730 & \textbf{1.397} & 1.708 && 1.195 & 1.229 & \textbf{1.194} \\
		0.4 & 2.436 & \textbf{1.529} & 2.297 && 1.460 & \textbf{1.128} & 1.410 && 1.110 & 1.181 & \textbf{1.108} \\
		0.5 & 1.911 & \textbf{1.041} & 1.691 && 1.263 & \textbf{1.033} & 1.196 && 1.058 & 1.154 & \textbf{1.053} \\
		0.6 & 1.491 & \textbf{1.013} & 1.246 && 1.124 & \textbf{1.030} & 1.069 && 1.029 & 1.135 & \textbf{1.024} \\
		0.7 & 1.196 & \textbf{1.013} & 1.045 && 1.049 & 1.028 & \textbf{1.019} && 1.017 & 1.114 & \textbf{1.014} \\
		0.8 & 1.050 & 1.012 & \textbf{1.008} && 1.016 & 1.027 & \textbf{1.008} && 1.013 & 1.091 & \textbf{1.012} \\
		0.9 & 1.009 & 1.012 & \textbf{1.005} && 1.007 & 1.026 & \textbf{1.006} && 1.012 & 1.073 & \textbf{1.011} \\
		1.0 & \textbf{1.004} & 1.012 & \textbf{1.004} && \textbf{1.006} & 1.026 & \textbf{1.006} && \textbf{1.011} & 1.065 & \textbf{1.011} \\
		\bottomrule
	\end{tabular}	
	\label{table:n=500}
\end{table}
%The average number of items per strip for $W = 5000$, 2500 and 1250 are 8.475, 4.310, and 2.165 respectively.

\begin{table}[h!]
	\centering
	\caption{Average solution quality $q$ for $n = 1000$.}
	\begin{tabular}{@{}c@{\hspace{15pt}}c@{\hspace{7pt}}c@{\hspace{5pt}}ccc@{\hspace{7pt}}c@{\hspace{5pt}}ccc@{\hspace{7pt}}c@{\hspace{5pt}}c@{}}\toprule
		& \multicolumn{3}{c}{$W = 5000$, $t = 115.534$} && \multicolumn{3}{c}{$W = 2500$, $t = 230.563$} && \multicolumn{3}{c}{$W = 1250$, $t = 460.623$} \\
		\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12}
		$\delta$ & MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$\\ \midrule
		0.0 & 8.657 & 8.657 & 8.657 && 4.338 & 4.338 & 4.338 && 2.171 & 2.171 & 2.171 \\
		0.1 & 5.173 & \textbf{4.842} & 5.140 && 2.636 & \textbf{2.481} & 2.643 && 1.511 & \textbf{1.467} & 1.511 \\
		0.2 & 3.976 & \textbf{3.462} & 3.952 && 2.093 & \textbf{1.858} & 2.102 && 1.318 & \textbf{1.311} & 1.319 \\
		0.3 & 3.047 & \textbf{2.350} & 3.013 && 1.698 & \textbf{1.409} & 1.688 && \textbf{1.183} & 1.221 & \textbf{1.183} \\
		0.4 & 2.374 & \textbf{1.520} & 2.266 && 1.426 & \textbf{1.131} & 1.384 && 1.097 & 1.171 & \textbf{1.096} \\
		0.5 & 1.847 & \textbf{1.026} & 1.642 && 1.230 & \textbf{1.030} & 1.170 && 1.044 & 1.144 & \textbf{1.040} \\
		0.6 & 1.433 & \textbf{1.012} & 1.203 && 1.099 & \textbf{1.027} & 1.050 && 1.020 & 1.124 & \textbf{1.015} \\
		0.7 & 1.155 & \textbf{1.012} & 1.026 && 1.034 & 1.026 & \textbf{1.012} && 1.012 & 1.102 & \textbf{1.009} \\
		0.8 & 1.032 & 1.012 & \textbf{1.006} && 1.011 & 1.025 & \textbf{1.007} && 1.009 & 1.078 & \textbf{1.008} \\
		0.9 & 1.007 & 1.012 & \textbf{1.004} && 1.007 & 1.024 & \textbf{1.006} && \textbf{1.008} & 1.059 & \textbf{1.008} \\
		1.0 & \textbf{1.004} & 1.012 & \textbf{1.004} && \textbf{1.006} & 1.024 & \textbf{1.006} && \textbf{1.008} & 1.054 & \textbf{1.008} \\
		\bottomrule
	\end{tabular}
	\label{table:n=1000}	
\end{table}
% The average number of items per strip for $W = 5000$, 2500 and 1250 are 8.621, 4.329 and 2.169 respectively.

\noindent Tables \ref{table:n=500} and \ref{table:n=1000} compare the results obtained from the three heuristics using the different values of $\delta$ and $W$, for $n = 500$ and 1000 respectively. Figures in bold indicate the best average solution quality for the given combination of parameters. We see that $q$ tends toward 1 as $\delta$ increases since the proportion of score widths meeting the vicinal sum constraint increases, allowing more items per strip\footnote{The average number of items per strip when $n = 500$ for $W = 5000$, 2500 and 1250 are 8.475, 4.310, and 2.165 respectively, and the average number of items per strip when $n = 1000$ for $W = 5000$, 2500 and 1250 are 8.621, 4.329, and 2.169 respectively.} and reducing the number of strips required. Note that when $\delta = 1.0$ MFFD and MFFD$^+$ have identical solution qualities, as the instances are equivalent to the original BPP ($\tau = 0$); hence they operate in the same fashion as the original FFD heuristic.

>>>>

\begin{definition}
	\label{defn:althamcycle}
	Let $G(V, B\cup R)$ be a simple, undirected graph where each edge is a member of one of two sets, $B$ or $R$. $G$ contains an alternating Hamiltonian cycle if there exists a Hamiltonian cycle such that successive edges alternate between sets $B$ and $R$.
\end{definition}

\noindent It is clear that an alternating Hamiltonian cycle on $G$, if one exists, correpsonds to a feasible solution $\mathcal{T}$, as each ``blue'' edge from $B$ represents a pair of values in $\mathcal{M}$, and each ``red'' edge from $R$ indicates the values that meet the vicinal sum constraint, and thus can be placed alongside one another feasibly. Also, note that every edge in $B$ must be present in the alternating Hamiltonian cycle. Consequently, the task can also be seen as finding a matching $R' \subseteq R$ of cardinality $n+1$ such that the edge sets $B$ and $R'$ form an alternating Hamiltonian cycle as described in Definition \ref{defn:althamcycle}.

The problem of finding an alternating Hamiltonian cycle on general graphs is NP-hard as it generalises the classical Hamiltonian cycle problem \cite{haggkvist1977}. However, the special structure of graphs derived from instances of the COP allows them to be solved in polynomial-time using the following method.

To find a suitable matching $R' \subseteq R$, the Maximum Cardinality Matching (MCM) algorithm is executed, which is based on \cite{becker2010} and also the earlier work of Mahadev and Peled \cite{mahadev1994}. First, take each vertex $v_1, v_2,..., v_{2n+2}$ in turn, and select the edge in $R$ connecting $v_i$ to the highest-indexed vertex $v_j$ that is not already incident to an edge in $R'$. Now, add this edge $(v_i, v_j)$ to $R'$, and proceed to the next vertex until all the vertices have been assessed. The two vertices incident to each matching edge in $R'$ are now referred to as being ``matched''. Similarly to partners, let $m : V \to V$ be a bijective function that assigns each vertex with its match, $m(v_i) = v_j$. Then, we can denote this matching set as $R' =\{(v_i, m(v_i)) : v_i \in V\}$.

During MCM, if a vertex $v_i$ is not adjacent to any other unmatched vertex except its partner $p(v_i)$ via a blue edge, the preceeding vertex $v_{i-1}$ can be ``rematched'', provided that 
\begin{enumerate*}[label={(\alph*)}]
	\item $v_i$ is not the first vertex; 
	\item the previous vertex $v_{i-1}$ has been matched; and
	\item $v_{i-1}$ and $p(v_i)$ are adjacent via a red edge in $R$.
\end{enumerate*}
Then, $v_i$ is matched with the vertex that is currently matched with $v_{i-1}$, and $v_{i-1}$ is rematched with $p(v_i)$. Due to the initial order of the vertices, MCM is guaranteed to produce a maximum cardinality matching.

Clearly, if $|R'| < n+1$ after MCM has completed, there are an insufficient number of red edges to form an alternating Hamiltonian cycle, and therefore no feasible solution exists for the given instance $\mathcal{M}$. Otherwise, $R'$ is a perfect matching of cardinality $n+1$ and the spanning subgraph $G'=(V, B \cup R')$ is a 2-regular graph, where each vertex $v_i \in V$ is incident to one blue edge and one red edge, as can be seen in Figure \ref{fig:mps}. $G'$ therefore consists of one or more cyclic components $C_1, C_2, ..., C_l$.

@@

\noindent In the case where $G'$ comprises one component, i.e. $l = 1$, then this alternating cycle is in fact Hamiltonian, and therefore specifies a solution to the COP. However, if $l > 1$, the components of $G'$ must be combined to form a single alternating Hamiltonian cycle. To do this, an additional procedure called Bridge Recognition (BR) is executed that selects suitable edges in $R\backslash R'$ to replace edges in $R'$ to connect the components of $G'$.

BR operates by first ordering the edges in $R'$ such that the lower-indexed vertices of each edge are in increasing order and the higher-indexed vertices are in decreasing order. Any edges that cannot be placed in such an order are then removed from this list. For instance, in the example in Figure \ref{fig:mps}, the edges would be sorted as follows: $(v_1, v_{14}), (v_2, v_{13}),(v_3, v_{12}),(v_4, v_{11}),(v_6, v_{10}),(v_7, v_8)$. Note that, since $v_5$ was not matched with the highest-indexed vertex possible during MCM, the edge $(v_5, v_9)$ does not adhere to the required structure of the list and is therefore omitted.

@@

\noindent To find these overlapping sets, a Modified Bridge Recognition (MBR) algorithm is used, which is based on BR. Firstly, a copy of the set $R_i$ with the highest cardinality\footnote{In the event of a tie, MBR selects the set with the lowest index.} generated by BR is created, called $R^{*}_1$, and is added to $\mathcal{R}^{*}$. Then, MBR takes the original sorted list used in BR and removes edges from the list that are already in $\mathcal{R}^{*}$. Similarly to BR, MBR proceeds through the list to find an edge that meets the following conditions:
\begin{enumerate*}[label={(\alph*)}]
	\item The lower-indexed vertex of the current edge and the higher-indexed vertex of the next edge in the list are adjacent via an edge in $R$; and
	\item the current edge and an edge in $\mathcal{R}^{*}$ are members of the same component of $G'$, and the next edge is a member of a component that does not have an edge in $\mathcal{R}^{*}$, or vice versa.
\end{enumerate*}
If both conditions hold, the current edge is added to a new set $R^{*}_2$ which is then added to $\mathcal{R}^{*}$. MBR continues to add suceeding edges to $R^{*}_2$, provided condition (a) holds and the suceeding edge is a member of a component that does not have an edge in $\mathcal{R}^{*}$, stopping when an edge that does not meet these conditions is reached. Then, if every component of $G'$ has an edge in one of the sets in $\mathcal{R}^{*}$, these sets are able to connect all the components of $G'$ together, and so MBR terminates. Otherwise, the edges in $R^{*}_2$ are removed from the sorted list, and MBR repeats the search for suitable edges to start a new set $R^{*}_3$. This procedure continues until either $\mathcal{R}^{*}$ contains overlapping sets that cover all components of $G'$, or until there are no more suitable edges in the list to start a new set. If MBR has produced a feasible collection of sets then the components of $G'$ can be merged to create an alternating Hamiltonian cycle by applying the connecting procedure above to every set in $\mathcal{R}^{*}$.

Using the instance illustrated in Figure \ref{fig:connect/cycle} as an example, the edges $(v_3, v_{12})$ and $(v_4, v_{11})$ are in the set $R_1$ formed by BR, as \begin{enumerate*}[label={(\alph*)}]
	\item $(v_3, v_{11}) \in R$, i.e. the lower-indexed vertex of the first edge is adjacent to the higher-indexed vertex of the next edge; and
	\item the edges are members of different components ($(v_3, v_{12}) \in C_1$ and $(v_4, v_{11}) \in C_2$ as shown previously in Figure \ref{fig:mps}).
\end{enumerate*}
Note that since $v_4$ and $v_{11}$ are adjacent, $v_4$ must also be adjacent to $v_{12}$, i.e. $(v_4, v_{12}) \in R$, since the value associated with $v_{12}$ is greater than or equal to the value associated with $v_{11}$. Then, because $|R_1| = l$, the edges $(v_3, v_{12})$ and $(v_4, v_{11})$ are removed from $R'$ and replaced by the edges $(v_3, v_{11})$ and $(v_4, v_{12})$ from $R_1$. Removing the dominating vertices and any incident edges results in an alternating Hamiltonian path, which corresponds to a feasible solution $\mathcal{T}$.

@@

\section{Experimental Results}
Benchmark instances currently do not exist for the SCSPP, and so artificial problem instances were produced to compare the performance of the three heuristics. For our experiments, we generated 1000 problem instances for $|\mathcal{I}| \in \{500, 100\}$. In each problem instance the items have varying widths $w_i \in [150,1000]$ and score widths $a_i, b_i \in [1, 70]$ to ensure that each item has exactly two score lines, all selected randomly from a uniform distribution. Strips of widths $W = 5000$, 2500 and 1250 were used to influence the number of items per strip. As the width of the strips decrease, the average number of items per strip also decreases, making the problem more constrained. Both the items and the strips have equal height of $H = 1$. Similarly to experiments performed in \cite{lewis2011}, we also introduced a parameter $\delta$ to denote the proportion of pairs of score widths from different items that meet the vicinal sum constraint, i.e. whose sum is greater than or equal to $\tau$. Values of $\delta$ from 0.0 to 1.0 were created by changing the value of $\tau$. Clearly, when $\delta = 0.0$, there are no items that can be packed together feasibly, and so $n$ strips will be required (one for each item), whereas if $\delta = 1.0$ all pairs of score widths meet the constraint, and the problem is equivalent to the BPP.

The heuristics were implemented in C++ and executed on a computer with an Intel Core i3-2120 3.30GHz processor. Our source code and all data is provided at \cite{hawa2018}. Since optimal solutions are not available, in our case solution quality $q$ is estimated by comparing each solution to the theoretical minimum, $q = |\mathcal{S}| / t$. For each heuristic, we calculated the average solution quality for every combination of $n$, $\delta$, and $W$ from 1000 instances. All individual trials were seen to complete in under 160ms.

\begin{table}[h!]
	\centering
	\caption{Average solution quality $q$ for $n = 500$.}
	\begin{tabular}{c@{\hspace{15pt}}c@{\hspace{7pt}}c@{\hspace{4pt}}ccc@{\hspace{7pt}}c@{\hspace{4pt}}ccc@{\hspace{7pt}}c@{\hspace{4pt}}c}\toprule
		& \multicolumn{3}{c}{$W = 5000$, $t = 58.039$} & \phantom{a} & \multicolumn{3}{c}{$W = 2500$, $t = 115.571$} & \phantom{a} & \multicolumn{3}{c}{$W = 1250$, $t = 230.648$} \\
		\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12}
		$\delta$ & MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$\\ \midrule
		0.0 & 8.618 & 8.618 & 8.618 && 4.328 & 4.328 & 4.327 && 2.169 & 2.169 & 2.169 \\
		0.1 & 5.214 & \textbf{4.842} & 5.161 && 2.659 & \textbf{2.477} & 2.657 && 1.515 & \textbf{1.479} & 1.515 \\
		0.2 & 4.031 & \textbf{3.459} & 3.976 && 2.121 & \textbf{1.847} & 2.118 && 1.326 & \textbf{1.318} & 1.326 \\
		0.3 & 3.111 & \textbf{2.348} & 3.038 && 1.730 & \textbf{1.397} & 1.708 && 1.195 & 1.229 & \textbf{1.194} \\
		0.4 & 2.436 & \textbf{1.529} & 2.297 && 1.460 & \textbf{1.128} & 1.410 && 1.110 & 1.181 & \textbf{1.108} \\
		0.5 & 1.911 & \textbf{1.041} & 1.691 && 1.263 & \textbf{1.033} & 1.196 && 1.058 & 1.154 & \textbf{1.053} \\
		0.6 & 1.491 & \textbf{1.013} & 1.246 && 1.124 & \textbf{1.030} & 1.069 && 1.029 & 1.135 & \textbf{1.024} \\
		0.7 & 1.196 & \textbf{1.013} & 1.045 && 1.049 & 1.028 & \textbf{1.019} && 1.017 & 1.114 & \textbf{1.014} \\
		0.8 & 1.050 & 1.012 & \textbf{1.008} && 1.016 & 1.027 & \textbf{1.008} && 1.013 & 1.091 & \textbf{1.012} \\
		0.9 & 1.009 & 1.012 & \textbf{1.005} && 1.007 & 1.026 & \textbf{1.006} && 1.012 & 1.073 & \textbf{1.011} \\
		1.0 & \textbf{1.004} & 1.012 & \textbf{1.004} && \textbf{1.006} & 1.026 & \textbf{1.006} && \textbf{1.011} & 1.065 & \textbf{1.011} \\
		\bottomrule
	\end{tabular}	
	\label{table:n=500}
\end{table}

\begin{table}[h!]
	\centering
	\caption{Average solution quality $q$ for $n = 1000$.}
	\begin{tabular}{c@{\hspace{15pt}}c@{\hspace{7pt}}c@{\hspace{4pt}}ccc@{\hspace{7pt}}c@{\hspace{4pt}}ccc@{\hspace{7pt}}c@{\hspace{4pt}}c}\toprule
		& \multicolumn{3}{c}{$W = 5000$, $t = 115.534$} & \phantom{a} & \multicolumn{3}{c}{$W = 2500$, $t = 230.563$} & \phantom{a} & \multicolumn{3}{c}{$W = 1250$, $t = 460.623$} \\
		\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12}
		$\delta$ & MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$\\ \midrule
		0.0 & 8.657 & 8.657 & 8.657 && 4.338 & 4.338 & 4.338 && 2.171 & 2.171 & 2.171 \\
		0.1 & 5.173 & \textbf{4.842} & 5.140 && 2.636 & \textbf{2.481} & 2.643 && 1.511 & \textbf{1.467} & 1.511 \\
		0.2 & 3.976 & \textbf{3.462} & 3.952 && 2.093 & \textbf{1.858} & 2.102 && 1.318 & \textbf{1.311} & 1.319 \\
		0.3 & 3.047 & \textbf{2.350} & 3.013 && 1.698 & \textbf{1.409} & 1.688 && \textbf{1.183} & 1.221 & \textbf{1.183} \\
		0.4 & 2.374 & \textbf{1.520} & 2.266 && 1.426 & \textbf{1.131} & 1.384 && 1.097 & 1.171 & \textbf{1.096} \\
		0.5 & 1.847 & \textbf{1.026} & 1.642 && 1.230 & \textbf{1.030} & 1.170 && 1.044 & 1.144 & \textbf{1.040} \\
		0.6 & 1.433 & \textbf{1.012} & 1.203 && 1.099 & \textbf{1.027} & 1.050 && 1.020 & 1.124 & \textbf{1.015} \\
		0.7 & 1.155 & \textbf{1.012} & 1.026 && 1.034 & 1.026 & \textbf{1.012} && 1.012 & 1.102 & \textbf{1.009} \\
		0.8 & 1.032 & 1.012 & \textbf{1.006} && 1.011 & 1.025 & \textbf{1.007} && 1.009 & 1.078 & \textbf{1.008} \\
		0.9 & 1.007 & 1.012 & \textbf{1.004} && 1.007 & 1.024 & \textbf{1.006} && \textbf{1.008} & 1.059 & \textbf{1.008} \\
		1.0 & \textbf{1.004} & 1.012 & \textbf{1.004} && \textbf{1.006} & 1.024 & \textbf{1.006} && \textbf{1.008} & 1.054 & \textbf{1.008} \\
		\bottomrule
	\end{tabular}
	\label{table:n=1000}	
\end{table}

\noindent Tables \ref{table:n=500} and \ref{table:n=1000} compare the results obtained from the three heuristics using the different values of $\delta$ and $W$, for $n = 500$ and 1000 respectively. Figures in bold indicate the best average solution quality for the given combination of parameters. We see that $q$ tends towards 1 as $\delta$ increases since the proportion of score widths meeting the vicinal sum constraint increases, consequently permissing more items to be packed on each individual strip\footnote{The average number of items per strip when $n = 500$ for $W = 5000$, 2500 and 1250 are 8.475, 4.310, and 2.165 respectively, and the average number of items per strip when $n = 1000$ for $W = 5000$, 2500 and 1250 are 8.621, 4.329, and 2.169 respectively.}, thus reducing the number of strips required. Note that when $\delta = 1.0$ MFFD and MFFD$^+$ have identical solution qualities, as the instances are equivalent to the original BPP ($\tau = 0$); hence they operate in the same fashion as the original FFD heuristic.
